---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lattice)
```

## Loading and prepocessing the data

Opening the data set:

```{r}
data <- read.csv2("~/R/RSTUDIO/COursera Data Science/Modul 5/Week2/activity.csv", header=TRUE, sep=",")
```
Convert date and time
```{r}
mydates <- as.Date(data$date, "%y-%m-%d")
```

## What is the mean total number of steps taken per day?

    ###1 - Histogram of total of steps per day


```{r}
steps_day <- data %>% 
    group_by(date) %>% 
    summarize(Steps=sum(steps, na.rm=TRUE))

head(steps_day)

hist(steps_day$Steps, main="Total steps taken each day", xlab="Number of Steps a day", ylab= "", col="blue")                 
```
    
   
    ###2 - Mean and Median of the total number of Steps per Day
    
```{r}
mean_steps <- mean(steps_day$Steps, na.rm=TRUE)

mean_steps

median_steps <- median(steps_day$Steps, na.rm=TRUE)

median_steps
```

## What is the average daily activity pattern?

    ###1 - Average number of steps in the 5 min interval
    
```{r}
steps_interval <- data %>%
  group_by(interval) %>%
  summarise(Steps=mean(steps, na.rm=TRUE))

head(steps_interval)

plot(steps_interval$interval, steps_interval$Steps, type="l", xlab="Interval", ylab="Average Steps per day", main="Average Steps per day in 5 m interval")
```
   
   
    ###2 - Interval with maximum number of steps
    
```{r}
Interval_max_steps <- steps_interval$interval[which.max(steps_interval$Steps)]
Interval_max_steps
```
**Interval with the maximum number of Steps:835**


## Imputing missing values

    ###1 - Total of Missing Values
    
```{r}
Total_rows_na <- sum(is.na(data))
Total_rows_na
```


      
      ###2, 3- Filling in all Missing Values and creating a new data set
```{r}
data_2 <- data
for (i in 1:nrow(data_2)) {
    if (is.na(data_2$steps[i])) {
        ndx <- which(data_2$interval[i] == steps_interval$interval)
        data_2$steps[i] <- steps_interval[ndx,]$Steps
    }
}
summary(data_2$steps)
```
      
      ### 4 - Making an Histogram of the Total number of steps each day as well as calculatin the Mean and Median
      
```{r}
steps_day2 <- data_2 %>% 
    group_by(date) %>% 
    summarize(Steps=sum(steps))

head(steps_day2)

hist(steps_day2$Steps, main="Total steps taken each day", xlab="Number of Steps a day", ylab= "", col="violet") 

mean_steps2 <- mean(steps_day2$Steps)

mean_steps2

median_steps2 <- median(steps_day2$Steps)

median_steps2
```


## Are there differences in activity patterns between weekdays and weekends?


      ###1 - Create variable with two levels (weekday and weekend)
      
```{r}
day <- weekdays(as.Date(data_2$date))
daytype<- vector()
for (i in 1:nrow(data_2)) {
    if (day[i] == "Saturday") {
        daytype[i] <- "Weekend"
    } else if (day[i] == "Sunday") {
        daytype[i] <- "Weekend"
    } else {
        daytype[i] <- "Weekday"
    }
}
data_2$daytype <- daytype
data_2$daytype<- factor(data_2$daytype)
```
      
      ###2 - Panel plot containing a time series plot of the 5m interval and the average number of steps taken, avereged across all weekdays or weekends
      
```{r}
Average <- aggregate(steps ~ interval + daytype, data=data, mean)
names(Average) <- c("interval", "daytype", "steps")

xyplot(steps ~ interval | daytype, Average, type = "l", layout = c(1, 2), xlab = "Interval", ylab = "Number of steps", main="Comparison of the average steps per day type") 
```


